Chasing \rainbows{} with the Habitable Worlds Observatory

### 2023-04-25 src/tex/ms.tex
[showyourwork] overleaf sync

% Authors are in alphabetical order:
\author[Sophia R. Vaughan et al.]{%
    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
    Sarah L. Casewell\textsuperscript{\afflink{5}},
    Nicolas B. Cowan\textsuperscript{\afflink{6}},
    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
    \newauthor
    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
    Matthew Kenworthy\textsuperscript{\afflink{7}},
    Johan Mazoyer\textsuperscript{\afflink{11}},
    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
    \newauthor
    Olivier Absil\textsuperscript{\afflink{13}},
    Lisa Altinier\textsuperscript{\afflink{14}},
    Pierre Baudoz\textsuperscript{\afflink{11}},
    Ruslan Belikov\textsuperscript{\afflink{15}},
    Alexis Bidot\textsuperscript{\afflink{16}},
    Markus J. Bonse\textsuperscript{\afflink{10}},
    \newauthor
    Bernhard Brandl\textsuperscript{\afflink{7}},
    Alexis Carlotti\textsuperscript{\afflink{16}},
    Elodie Choquet\textsuperscript{\afflink{14}},
    Niyati Desai\textsuperscript{\afflink{17}},
    Kevin Fogarty\textsuperscript{\afflink{15}},
    Jules Fowler\textsuperscript{\afflink{18}},
    \newauthor
    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
    \newauthor
    Adrien Hours\textsuperscript{\afflink{16}},
    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
    Elina Kleisioti\textsuperscript{\afflink{7}},
    Lorenzo König\textsuperscript{\afflink{13}},
    Mariya Krasteva\textsuperscript{\afflink{27}},
    \newauthor
    Iva Laginja\textsuperscript{\afflink{11}},
    Rico Landman\textsuperscript{\afflink{7}},
    Lucie Leboulleux\textsuperscript{\afflink{16}},
    David Mouillet\textsuperscript{\afflink{16}},
    Mamadou N’Diaye\textsuperscript{\afflink{28}},
    \newauthor
    Emiel H. Por\textsuperscript{\afflink{29}},
    Laurent Pueyo\textsuperscript{\afflink{29}},
    Frans Snik\textsuperscript{\afflink{7}},
    Daphne M. Stam\textsuperscript{\afflink{30}},
    Victor Trees\textsuperscript{\afflink{31},\afflink{32}},
    Dirk van Dam\textsuperscript{\afflink{7}},
    \newauthor
    Kyle van Gorkom\textsuperscript{\afflink{21}},
    Maaike van Kooten\textsuperscript{\afflink{33}}
    \newauthor \\%
    Affiliations are listed at the end of the paper
    % Timmy: I checked MNRAS and it seems that this is how they handle very long author lists. See, for example:
    % https://academic.oup.com/mnras/article/498/2/2354/5900562
}
\date{Draft version: \today}

@@

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search for biosignatures. 
%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal. After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties. 
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering phenomena. 
%
Direct imaging missions are optimised to detect planets near quadrature, which may obscure some of the phases at which these features are strongest. 
%
The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. 
%
Planets that orbit close to edge-on have accessible phase angles limited by obscuration by the coronagraph. 
%
We use the list of target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
%
We find that exo-Earths in systems listed in this Habitable Worlds Observatory catalog will have a detectable Rayleigh scattering peak. 
%
The glint signature at planetary phase angles of $\sim\qtyrange{50}{70}{\degree}$ 
%\sim$110--130$^\circ$ 
would be accessible to the Habitable Worlds Observatory in $\sim\num{14}$ systems, and the rainbow of water clouds at phases of \qtyrange{140}{160}{\degree} would be accessible in $\sim\num{42}$ systems, assuming a \qty{62}{\mas} inner working angle (\IWA{}; $3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
%
Increasing the \IWA{} to $2\,\lambda/D$ increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
%
The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA{}.  \end{abstract}

@@

In unpolarized light {\color{Green}{do you mean polarized light?}}, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
Lambertian reflecting planet.
%
{\color{red}{check the sentence:}}
While the glory and rainbow are technically visible in any phase curve
in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
%
The rainbow angle depends on the refractive index of the scatterers and thus depends on the composition and the wavelength. 
The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
%
Because the rainbow feature is more narrow than the glint, it would 
allow for a more exact determination of the particle composition.
%
Note that the scattering by clouds that consist of solid particles, 
such as ice crystals, will produce other optical features, such as halo's, particularly at small scattering angles \citep{bailey2007, hansentravis1974, karalidi2012rainbow}.
%

@@

%--------------------------------------------------------------------------
\begin{figure*}%[t]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/accessible_phase_angles.png}  
    \caption{
        \todo{shouldn't lambda/D for 600nm give 21, 41, 62, 83 mas?} 
        The cumulative distributions of the most extreme scattering phase angles accessible for different \IWA{} and for randomly inclined, circular orbits
        (solid lines) and randomly inclined, elliptical orbits (dashed lines).
        %
        The top $x$-axis indicates the minimum and the 
        bottom $x$-axis the maximum accessible scattering phase angle.
        %
        These angles are symmetric about quadrature (90 degrees).
        %
        The $y$-axis on the left indicates the number of planetary systems 
        divided by the number of Monte Carlo samples and is thus normalised to 
        the number of systems on the target list.
        %
        The $y$-axis on the right indicates the number of systems, assuming \qty{24}{\percent} of them have an Earth-like planet in their HZ.
    }
    \label{fig:betaallofit}
    \script{plot_n_versus_phase.py}
\end{figure*}
%--------------------------------------------------------------------------

>>>>

% Authors are in alphabetical order:
\author[Sophia R. Vaughan et al.]{%
    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
    Sarah L. Casewell\textsuperscript{\afflink{5}},
    Nicolas B. Cowan\textsuperscript{\afflink{6}},
    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
    \newauthor
    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
    Matthew A.  Kenworthy\textsuperscript{\afflink{7}},
    Johan Mazoyer\textsuperscript{\afflink{11}},
    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
    \newauthor
    Olivier Absil\textsuperscript{\afflink{13}},
    Lisa Altinier\textsuperscript{\afflink{14}},
    Pierre Baudoz\textsuperscript{\afflink{11}},
    Ruslan Belikov\textsuperscript{\afflink{15}},
    Alexis Bidot\textsuperscript{\afflink{16}},
    Markus J. Bonse\textsuperscript{\afflink{10}},
    \newauthor
    Bernhard Brandl\textsuperscript{\afflink{7}},
    Alexis Carlotti\textsuperscript{\afflink{16}},
    Elodie Choquet\textsuperscript{\afflink{14}},
    Niyati Desai\textsuperscript{\afflink{17}},
    Kevin Fogarty\textsuperscript{\afflink{15}},
    Jules Fowler\textsuperscript{\afflink{18}},
    \newauthor
    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
    \newauthor
    Adrien Hours\textsuperscript{\afflink{16}},
    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
    Elina Kleisioti\textsuperscript{\afflink{7}},
    Lorenzo König\textsuperscript{\afflink{13}},
    Mariya Krasteva\textsuperscript{\afflink{27}},
    \newauthor
    Iva Laginja\textsuperscript{\afflink{11}},
    Rico Landman\textsuperscript{\afflink{7}},
    Lucie Leboulleux\textsuperscript{\afflink{16}},
    David Mouillet\textsuperscript{\afflink{16}},
    Mamadou N’Diaye\textsuperscript{\afflink{28}},
    \newauthor
    Emiel H. Por\textsuperscript{\afflink{29}},
    Laurent Pueyo\textsuperscript{\afflink{29}},
    Frans Snik\textsuperscript{\afflink{7}},
    Daphne M. Stam\textsuperscript{\afflink{30}},
    Victor Trees\textsuperscript{\afflink{31},\afflink{32}},
    Dirk van Dam\textsuperscript{\afflink{7}},
    \newauthor
    Kyle van Gorkom\textsuperscript{\afflink{21}},
    Maaike van Kooten\textsuperscript{\afflink{33}}
    \newauthor \\%
    Affiliations are listed at the end of the paper
    % Timmy: I checked MNRAS and it seems that this is how they handle very long author lists. See, for example:
    % https://academic.oup.com/mnras/article/498/2/2354/5900562
}
\date{Draft version: \today}

@@

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory (\hwo), a coronagraphic mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures. 
%Numerous instrumental and observational challenges have still to be overcome to design a coronagraphic mission with that goal. After a research phase aiming at the detection of a few tens of exo-Earth in the habitable zone, the Habitable Worlds Observatory will use its spectroscopic and polarimetric capabilities to probe these planets' atmosphere and surface properties. 
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase variations can show the presence of specularly reflecting oceans on exoplanets, in addition to being sensitive to rainbows and other cloud scattering phenomena. 
%
Direct imaging missions are optimised to detect planets near quadrature, and the coronagraph may obscure the phases at which these features are strongest. 
%
The range of scattering phases accessible for an exoplanet can be limited by its orbital inclination or the coronagraph's inner working angle. 
%
Planets that orbit close to edge-on have accessible phase angles limited by coronagraph obscuration. 
%
We use the list of 160 target stars for the Habitable Worlds Observatory to estimate the number of exo-Earths that could be searched for non-Lambertian scattering phenomena. 
%
We find that any exo-Earths in systems listed in this \hwo\ catalog will have a detectable Rayleigh scattering peak. 
%
The glint signature at planetary phase angles of $\sim\qtyrange{50}{70}{\degree}$ 
%\sim$110--130$^\circ$ 
would be accessible to the Habitable Worlds Observatory in $\sim\num{14}$ systems, and the rainbow of water clouds at phases of \qtyrange{140}{160}{\degree} would be accessible in $\sim\num{42}$ systems, assuming a \qty{62}{\mas} inner working angle (\IWA{}; $3\,\lambda/D$ for a \qty{6}{\meter} telescope at \qty{600}{\nano\meter}).
%
Increasing the \IWA{} to $2\,\lambda/D$ increases the number of systems with detectable rainbows and glints by factors of roughly 2 and 3, respectively.
%
The number of systems in which an exo-Earth could be searched for these scattering phenomena increases with orbital eccentricity and scales approximately inversely with \IWA{}.  \end{abstract}

@@

In unpolarized light, the scattering by cloud particles can dramatically change the shape of a planet's phase curve compared to that of a 
Lambertian reflecting planet.
%
{\color{red}{check the sentence:}}
While the glory and rainbow are technically visible in any phase curve
in the presence of spherical cloud and/or haze particles, their strongly polarized nature coupled with the suppression of signals from other effects at gibbous and full phases in polarized light make them far easier to characterize in polarized light (both in the degree of polarization and in full Stokes measurements) \citep{karalidi2011, stam2008, treesandstam2019}.
%
The rainbow angle depends on the refractive index of the scatterers and thus depends on the composition and the wavelength. 
The wavelength dependence itself also depends on the particles' size distribution, and the size distribution affects the angular shape of the rainbow \citep{karalidi2011}
%
Because the rainbow feature is more narrow than the glint, it would 
allow for a more exact determination of the particle composition.
%
Note that the scattering by clouds that consist of solid particles, 
such as ice crystals, will produce other optical features, such as halo's, particularly at small scattering angles \citep{bailey2007, hansentravis1974, karalidi2012rainbow}.
%

@@

%--------------------------------------------------------------------------
\begin{figure*}%[t]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/accessible_phase_angles.png}  
    \caption{
        The cumulative distributions of the most extreme scattering phase angles accessible for different \IWA{} and for randomly inclined, circular orbits
        (solid lines) and randomly inclined, elliptical orbits (dashed lines).
        %
        The top $x$-axis indicates the minimum and the 
        bottom $x$-axis the maximum accessible scattering phase angle.
        %
        These angles are symmetric about quadrature (90 degrees).
        %
        The $y$-axis on the left indicates the number of planetary systems 
        divided by the number of Monte Carlo samples and is thus normalised to 
        the number of systems on the target list.
        %
        The $y$-axis on the right indicates the number of systems, assuming \qty{24}{\percent} of them have an Earth-like planet in their HZ.
    }
    \label{fig:betaallofit}
    \script{plot_n_versus_phase.py}
\end{figure*}
%--------------------------------------------------------------------------
